---
title: "Data_Analytics_Assignment_W21013663"
author: "Will Chan"
format: html
editor: visual
---

## 1. Install necessary packages

```{r}

#Command line if neccessary - remove the hashtag # before each install line to install packages only if neccessary.
# install.packages("tidyverse")
# install.packages("plotly")
# install.packages("ggplot2")
# install.packages("dplyr")

library(tidyverse)
library(plotly)
library(ggplot2)
library(dplyr)


```

## 2. Load in raw data sets and view

This data is from the Official for National Statistics website. This data set originally came with 3 tables:

-Table_1 shows data on E-cigarette use, by sex and age, England, 2014 to 2023

-Table_2 shows data on E-cigarette use, by sex and cigarette smoking status, England, 2014 to 2023

-Table_3 shows data on Confidence intervals for the adult e-cigarette estimates, England, 2014 to 2023

For this study, I will mainly focus on Table_1 as the age is necessary.

```{r}
Table_1_raw_data <-read.csv("https://raw.githubusercontent.com/WillnWill/Data_analytics_ecigarette-analysis/main/ecigaretteuseinengland1.csv")

View(Table_1_raw_data)

```

## 

## 2. Clean and separate Table_1 to show daily users of all gender.

```{r}
#1.Save the dataset
Table_1_DailyUsers_AllGender <- read.csv("https://raw.githubusercontent.com/WillnWill/Data_analytics_ecigarette-analysis/main/ecigaretteuseinengland1.csv")

#2.Convert it to a tibbles
Table_1_DailyUsers_AllGender <- as_tibble(Table_1_DailyUsers_AllGender) 


#3.Remove collumn
Table_1_DailyUsers_AllGender <- Table_1_DailyUsers_AllGender[, -c(3:14)]



#4.Remove unneccesary cells
Table_1_DailyUsers_AllGender <- Table_1_DailyUsers_AllGender[-c(1,2,3,4,5,6,7,8,9,10,11,13,14,15,16,17,19,20,21,22,23,25,26,27,28,29,31,32,33,34,35,36,37,38,40,41,42,43,44,46,47,48,49,50,51,52,53,55,56,57,58,59,61,62,63,64,65,67,68,69,70,71,73,74,75,76,77), ]



#5.Ensure headers are correctly positioned and labeled
colnames(Table_1_DailyUsers_AllGender) <- c(
  "Year",
  "Percentage of users",
  "All_16_to_24",
  "All_25_to_34",
  "All_35_to_49",
  "All_50_to_59",
  "All_60_and_over",
  "All_16_and_over"
)



#6.Remove all NA's

Table_1_DailyUsers_AllGender_long <- Table_1_DailyUsers_AllGender

Table_1_DailyUsers_AllGender_long <- na.omit(Table_1_DailyUsers_AllGender_long)

Table_1_DailyUsers_AllGender <- Table_1_DailyUsers_AllGender %>%
  select(where(~ !all(is.na(.))))

#7.Turn values numeric that start with All_
Table_1_DailyUsers_AllGender <- Table_1_DailyUsers_AllGender %>%
  mutate(across(starts_with("All_"), as.numeric))

Table_1_DailyUsers_AllGender <- Table_1_DailyUsers_AllGender %>%
  mutate(Year = as.numeric(Year))


#8.Show structure of the Table_1 daily users.
str(Table_1_DailyUsers_AllGender)

#9. Display the first few rows of each dataset in the console
head(Table_1_DailyUsers_AllGender)


#10. View the cleansed data on another file
View(Table_1_DailyUsers_AllGender)

```

## 3. Add a regression model for Table_1_DailyUsers_AllGender

```{r}

#1.Ensure that the year starts at 2014, not 0.
Table_1_DailyUsers_AllGender$Year_starting_2014 <- Table_1_DailyUsers_AllGender$Year - 2014

#2.Reshape the data to combine the 2 age groups.
data_long <- Table_1_DailyUsers_AllGender %>%
  pivot_longer(
    cols = c( All_16_to_24, All_35_to_49 ),
    names_to = "Age_Group",
    values_to = "E_Cigarette_Use"
  )


#3.Apply the multiple linear regression model
model <- lm( E_Cigarette_Use ~ Year_starting_2014 * Age_Group, data = data_long )

summary(model)






```

## 

## 5. Plot a regression model for Table_1_DailyUsers_AllGender to see e-cigarette use for both age groups.

```{r}

ggplot(data_long, aes(x = Year, y = E_Cigarette_Use, color = Age_Group)) +
  geom_point(size = 2) +  
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +  
  scale_color_manual(values = c("All_16_to_24" = "Darkgreen", "All_35_to_49" = "Darkorange")) +
  labs(
    title = "E-cigarette Use by Age Group (2014–2023)",
    x = "Years",
    y = "E-cigarette Use (%)",
    color = "Age Group"
  ) +
  scale_x_continuous(breaks = seq(2014, 2023, 1))


```

## 

## [See if gender affects the e-cigarette use.]{.underline}

This script below will analyse if gender affects the e-cigarette use. For this analysis, we will be focusing on the younger people (age 16 to 24).

## 1. Clean and separate Table_1 to have men e-cigarette use and female e-cigarette use in separate datasets.

```{r}

###FOR MALE GENDER DATASET CLEANSING *****************************************

#1.Save the dataset
Table_1_DailyUsers_MaleGender <- read.csv("https://raw.githubusercontent.com/WillnWill/Data_analytics_ecigarette-analysis/main/ecigaretteuseinengland1.csv")

#2.Convert it to a tibbles
Table_1_DailyUsers_MaleGender <- as_tibble(Table_1_DailyUsers_MaleGender) 

#3.Remove collumn
Table_1_DailyUsers_MaleGender <- Table_1_DailyUsers_MaleGender[, -c(9:21)]

#4.Remove unneccesary cells
Table_1_DailyUsers_MaleGender <- Table_1_DailyUsers_MaleGender[-c(1,2,3,4,5,6,7,8,9,10,11,13,14,15,16,17,19,20,21,22,23,25,26,27,28,29,31,32,33,34,35,36,37,38,40,41,42,43,44,46,47,48,49,50,51,52,53,55,56,57,58,59,61,62,63,64,65,67,68,69,70,71,73,74,75,76,77), ]

#5.Ensure headings are correctly positioned and labelled.
colnames(Table_1_DailyUsers_MaleGender) <- c(
  "Year",
  "Percentage of users",
  "All_16_to_24",
  "All_25_to_34",
  "All_35_to_49",
  "All_50_to_59",
  "All_60_and_over",
  "All_16_and_over"
)

#6.Remove all NA's
Table_1_DailyUsers_MaleGender <- Table_1_DailyUsers_MaleGender %>%
  select(where(~ !all(is.na(.))))  

Table_1_DailyUsers_MaleGender_long <- Table_1_DailyUsers_MaleGender

Table_1_DailyUsers_MaleGender <- na.omit(Table_1_DailyUsers_MaleGender_long)


#7.Turn values numeric
Table_1_DailyUsers_MaleGender <- Table_1_DailyUsers_MaleGender %>%
  mutate(across(starts_with("Men_"), as.numeric))

Table_1_DailyUsers_MaleGender <- Table_1_DailyUsers_MaleGender %>%
  mutate(Year = as.numeric(Year))

#8.Show structure of the Table_1 daily users.
str(Table_1_DailyUsers_MaleGender)

#9. Display the first few rows of each dataset in the console
head(Table_1_DailyUsers_MaleGender)

#10. View the cleansed data on another file
View(Table_1_DailyUsers_MaleGender)

#END OF MALE GENDER DATASET CLEANSING **********************************





#FOR FEMALE GENDER DATASET CLEANSING ******************************************

#1.Save the dataset
Table_1_DailyUsers_FemaleGender <- read.csv("https://raw.githubusercontent.com/WillnWill/Data_analytics_ecigarette-analysis/main/ecigaretteuseinengland1.csv")

#2.Convert it to a tibbles
Table_1_DailyUsers_FemaleGenderr <- as_tibble(Table_1_DailyUsers_FemaleGender) 


#3.Remove collumn
Table_1_DailyUsers_FemaleGender <- Table_1_DailyUsers_FemaleGender[, -c(3:8, 15:20)]


#4.Remove unneccesary cells
Table_1_DailyUsers_FemaleGender <- Table_1_DailyUsers_FemaleGender[-c(1,2,3,4,5,6,7,8,9,10,11,13,14,15,16,17,19,20,21,22,23,25,26,27,28,29,31,32,33,34,35,36,37,38,40,41,42,43,44,46,47,48,49,50,51,52,53,55,56,57,58,59,61,62,63,64,65,67,68,69,70,71,73,74,75,76,77), ]


#5.Ensure headings are correctly positioned and labelled.
colnames(Table_1_DailyUsers_FemaleGender) <- c(
  "Year",
  "Percentage of users",
  "All_16_to_24",
  "All_25_to_34",
  "All_35_to_49",
  "All_50_to_59",
  "All_60_and_over",
  "All_16_and_over"
)


#6.Remove all NA's
Table_1_DailyUsers_FemaleGender <- Table_1_DailyUsers_FemaleGender %>%
  select(where(~ !all(is.na(.))))  

Table_1_DailyUsers_FemaleGender_long <- Table_1_DailyUsers_FemaleGender

Table_1_DailyUsers_FemaleGender <- na.omit(Table_1_DailyUsers_FemaleGender_long)


#7.Turn values numeric
Table_1_DailyUsers_FemaleGender <- Table_1_DailyUsers_FemaleGender %>%
  mutate(across(starts_with("Women_"), as.numeric))

Table_1_DailyUsers_FemaleGender <- Table_1_DailyUsers_FemaleGender %>%
  mutate(Year = as.numeric(Year))

#8.Show structure of the Table_1 daily users.
str(Table_1_DailyUsers_FemaleGender)

#9. Display the first few rows of each dataset in the console
head(Table_1_DailyUsers_FemaleGender)

#10. View the cleansed data on another file
View(Table_1_DailyUsers_FemaleGender)

#END OF WOMEN DATASET CLEANSING *****************************************
```

## 2. Combine Table_1_DailyUsers_MaleGender and Table_1_DailyUsers_FemaleGender, focusing on age group 16 to 24.

```{r}

#1. Add Gender column to each dataset
Table_1_DailyUsers_MaleGender$Gender <- "Male"
Table_1_DailyUsers_FemaleGender$Gender <- "Female"

#2. Combine the two datasets into one
Table_1_DailyUsers_MixedGender <- rbind(Table_1_DailyUsers_MaleGender, Table_1_DailyUsers_FemaleGender)

#3. Ensure Gender is a factor (we need this for the comparison)
Table_1_DailyUsers_MixedGender$Gender <- as.factor(Table_1_DailyUsers_MixedGender$Gender)


#4.Convert it to a tibbles
Table_1_DailyUsers_MixedGender <- as_tibble(Table_1_DailyUsers_MixedGender) 

Table_1_DailyUsers_MixedGender$Year_starting_2014 <- Table_1_DailyUsers_MixedGender$Year - 2014

#5. Change the dataset to a long format
data_long <- Table_1_DailyUsers_MixedGender %>%
  pivot_longer(
    cols = starts_with("All_"),  # Columns for age groups
    names_to = "Age_Group",
    values_to = "E_Cigarette_Use"
  )


#6. Check the structure of the combined dataset.
str(Table_1_DailyUsers_MixedGender)

#7.View the combined dataset using converted tibbles dataset
View(Table_1_DailyUsers_MixedGender)


```

## 3. Apply a regression model

```{r}

model2 <- lm(E_Cigarette_Use ~ Year_starting_2014 * Age_Group * Gender, data = data_long)

summary(model2)
```

## 
